---
title: "Coursera Machine Learning Project"
author: "erwindijk"
date: "Sunday, December 21, 2014"
output: html_document
---

### Dataset
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
More information is available from the website here: 
http://groupware.les.inf.puc-rio.br/har  (see the section on the Weight Lifting Exercise Dataset).

### Project Goal
In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 
The goal is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. Any of the other variables can be used to predict with. Create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

### Classes
* Class A - Exercise performed 100% according to the specified specifications.
* Class B - Throwing the elbows to the front.
* Class C - Lifting the dumbbell only halfway.
* Class D - Lowering the dumbbell only halfway
* Class E - Throwing the hips to the front.

## Start analysis

### Data Processing = Load data and selection of subset 
```{r}
setwd("C:/Users/nl25484/Documents/R-projects/MachineLearning")
train <- read.table("pml-training.csv", head = TRUE, sep = ",")
test <- read.table("pml-testing.csv", head = TRUE, sep = ",")

```


Select variables to use in the analysis: 
* “classe”, which is going to be predicted, "problem" in the test set and 
* all variables which begin with “roll”, “pitch”, “yaw”, “total_accel”,“gyros”,“accel” and “magnet”. 
* Our new training set consists 52 variables and one "classe"" or "problem"" column
```{r}
names<-names(train)
subsetnames <- grep("^roll_|^pitch|^yaw_|^total_accel|^gyros_|^accel_|^magnet|classe",names,value=T)
pml_train <- subset(train,select=subsetnames)

names <- names(test)
subsetnames <- grep("^roll_|^pitch|^yaw_|^total_accel|^gyros_|^accel_|^magnet|classe|^problem",names,value=T)
pml_test<-subset(test,select=subsetnames)

names(pml_train)
```

### Spliting pml_train into a training and a testing set, use pml_test later  
```
library(caret)
set.seed(2134)
inTrain<-createDataPartition(pml_train$classe,p=0.6,list=FALSE)

train_set <- pml_train[inTrain,]
test_set <-pml_train[-inTrain,]
```

### PCA Pre-processing and RF Random Forext Modelling
```{r}
library(caret)

preProc <- preProcess(train_set[-53],method="pca",pcaComp=14)
trainPC <- predict(preProc,train_set[-53])
modelFit <- train(train_set$classe ~ .,method="rf",data=trainPC)
confusionMatrix(train_set$classe,predict(modelFit,trainPC))
```

### Apply to test_set
```{r}
testPC <- predict(preProc,test_set[-53])
confusionMatrix(test_set$classe,predict(modelFit,testPC))
```

### Results
The predicted values in pml-test set are the following
```{r}
pml_testPC<-predict(preProc,pml_test[-53])
predicted<-predict(modelFit,newdata=pml_testPC)
predicted
```


### Write files according to instructions
```{r}
answers = c("B", "A", "B", "A", "A", "E", "D", "B", "A", "A", 
            "B", "C", "B", "A", "E", "E", "A", "B", "B", "B") 

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

# then create a folder where you want the files to be written. Set that to be your working directory and run:
pml_write_files(predicted)

```

